\documentclass[12pt]{book}
\usepackage{../../thesis}

\begin{document}
\section{Conjugacy}
from Sternberg chap4
Iterations related by conjugation exhibit similar behaviors. From our perspective, they are pretty much the same.
Here are the things shared among mappings that are in the same class.
\begin{itemize}
  \item Topological transitivity
  \item Periodic points
\end{itemize}
\begin{proposition}
  Suppose that $g\circ h = h\circ f$ where $h$ is continuous and surjective. If $f$ is transitive,
  then it follows that $g$ is transitive.
\end{proposition}

It follows from the last two items in the list that mappings in the same conjugacy class are chaotic.
\begin{proposition}
  Let $h: X\rar Y$ be a surjective and continuous mapping. Suppose we have $h\circ f = g\circ f$. If $f: X\rar X$ is chaotic,
  then $g$ is chaotic.
\end{proposition}

\subsection{Conjugacy of maps}
These happen to be all conjugate to each other.
\begin{itemize}
  \item Logistic map ($L_\mu(x) = \mu x(1-x)$)
  \item Quadratic map ($Q_c(x) = x^2 + c$)
  \item Tent map 
    \begin{equation*}
      T(x) = 
      \begin{cases}
        2x & 0 \leq x \leq \frac{1}{2}      \\
        2 - 2x & \frac{1}{2} \leq x \leq 1
      \end{cases}
    \end{equation*}
  \item Reverse tent map ($V(x) = 2|x| - 2$)
  \item all quadratic maps ($p(x) = ax^2 + bx + d$)
  \item Sawtooth transformation 
    \begin{equation*}
      S(x) = 
      \begin{cases}
        2x     & 0 \leq x \leq \frac{1}{2}      \\
        2x - 1 & \frac{1}{2} \leq x \leq 1
      \end{cases}
    \end{equation*}
  \item Shift map ($Sh: a_1a_2a_3\ldots \mapsto a_2a_3a_4\ldots$)
  \item Doubling map ()
\end{itemize}

\begin{proposition}
  (Logistic Map and a quadratic map)
  Any quadratic map $f = ax^2 + bx + d$ is conjugate to 
\end{proposition}

\begin{proposition}
  (The doubling map and $Q_{-2}$)
  Let $h: S \rar [-2,2]$
  \begin{equation*}
    h(\theta) = 2\cos\theta.
  \end{equation*}
  $h$ is clearly surjective and continuous.
  The following equality establishes the conjugacy.
  \begin{equation*}
    h(D(\theta)) = 2\cos2\theta = 2(2\cos^2\theta - 1) = (2\cos\theta)^2 - 2 = Q_{-2}(h(\theta)).
  \end{equation*}
\end{proposition}

\begin{proposition}
  (The shift map and the sawtooth transformation)
  First we need to reformulate the sawtooth transformation in terms of the binary representation.
  It turns out to be easy. It's just a multiplication of a sequence of bits by 2, just like any computer scientist knows how to do:
  \begin{equation*}
    S: .a_1a_2a_3\ldots \mapsto .a_2a_3a_4\ldots
  \end{equation*}
  Then, let $h: X \rar [0,1)$ be a trivial transformation from a binary sequence into a binary representation of a number
    \begin{equation*}
      h: a_1a_2a_3\ldots \mapsto .a_1a_2a_3\ldots
    \end{equation*}
    (note the period to indicate fraction).
  Thus conjugacy is established
  \begin{equation*}
    S \circ h = h \circ Sh
  \end{equation*}
  \end{proposition}

  \begin{proposition}
    (The sawtooth transformation and the shift map)
    Let $h = T$, the tent map itself. We have four cases to verify:
    \begin{enumerate}[$(i)$]
      \item $[0,1/4)$
      \item $[1/4, 1/2)$
      \item $[1/2, 3/4)$
      \item $[3/4, 1]$
    \end{enumerate}
    Computation is straightforward but tedious. One can easily verify that
    for each case $T\circ T = T\circ S$ holds.
  \end{proposition}

\bibliographystyle{../../pjgsm}
\bibliography{../../bibliography/thesis}


\end{document}
