\documentclass[11pt]{article}
\usepackage{thesis}

\begin{document}
\section{Lyapunov Exponents}
Positive Lyapunov exponent.\citet{kantz-schreiber} 
The technique of Lyapunov exponents was developed by Alexander M. Lyapunov to study stability of solutions for systems of differential equations.

Lyapunov exponents are sometimes called characteristic exponents.
Lai-Sang Young gave a lower bound of the Hausdorff dimension in terms of the Lyapunov numbers (log of Lyapunov exponents).


\subsection{In One Dimension}
\begin{definition}
  (Lyapunov number)
  Given a map $F$ and a $p-$periodic point $x_p$, the Lyapunov number of the point $N(x_p)$ is
  defined to be the spectral radius of the Jacobian matrix of the map.
  \begin{equation*}
    N(x_p) = \left[ \rho\left( F^p(x_p) \right) \right]^{1/p}
  \end{equation*}
  \label{def:lyapnum}
\end{definition}

\begin{definition}
  (Lyapunov Number, alternative)
  Let
  \begin{equation*}
    O(x_1) = \set{x_1, x_2, x_3, \ldots}
  \end{equation*}
  and define matrices
  \begin{equation*}
    A_1 = \frac{d}{dx}F(x_1),\; A_2 = \frac{d}{dx}F(x_2),\; \ldots,\; A_n = \frac{d}{dx}F(x_n).
  \end{equation*}
  Then consider their spectral radii
  \begin{equation*}
    \rho_1(x_1), \rho_2(x_1),\; \ldots,\; \rho_n(x_1).
  \end{equation*}
  Then the alternate version of the Lyapunov number $N(x_1)$ of $O(x_1)$ is defined as 
  \begin{equation*}
    N(x_1) = \lim\limits_{n \to \infty} (\rho_n(x_1))^{1/n}.
  \end{equation*}
\end{definition}
The alternative definition can be shown to be equivalent to the first definition.

Lyapunov exponent is computationally preferable to the Lyapunov number.
\begin{definition}
  (Lyapunov exponent)
  \begin{equation*}
    \Lambda(x_1) = \log N(x_1).
  \end{equation*}
  In particular,
  \begin{equation*}
    \Lambda(x_1) = \lim\limits_{n \to \infty} \paren*{\log\abs{F'(x_n)} + \log\abs{F'(x_{n-1})} + \cdots + \log\abs{F'(x_1)}}
  \end{equation*}
\end{definition}

In some cases, this yet another alternate definition is more convenient.
This version ignores transient states.
\begin{definition}
  (Lyapunov Number, yet another alternative)
  \begin{equation*}
    N(x_1, k) = N(x_k)
  \end{equation*}
  Then the Lyapunov number is defined to be
  \begin{equation*}
    N(x_1) = \lim\limits_{k \to \infty} N(x_1,k).
  \end{equation*}
\end{definition}

It can be shown that the Lyapunov exponent of a map is (almost everywhere) equivalent to the Lyapunov exponent of any orbit of a map (Lyapunov exponent is independent of orbit).


\subsection{Lyapunov Exponents and Conjugacy}
Lyapunov number of one-dimensional dynamical systems that are conjugate is the same
as long as $h'(x) \neq 0$ and
\begin{equation*}
  \lim\limits_{n\to \infty} \frac{\log\abs{h'(x_n)}}{n} = 0.
\end{equation*}

By the definition of conjugacy,
\begin{equation*}
  h'(x_{n+1})F'(x_n)\ldots F'(x_1) = G'(y_n)\ldots G'(y_1)h'(x_1).
\end{equation*}
Then 
\begin{align*}
  \Lambda(x_1) &= \lim\limits_{n\to \infty} \frac{1}{n}\paren{\log\abs{F'(x_n)} + \cdots + \log\abs{F'(x_n)} } \\
  &= \lim\limits_{n\to \infty} \frac{1}{n}\left(\log\abs{F'(x_n)} + \cdots + \log\abs{F'(x_n)} + \log\abs{h'(x_{n+1})}\right) \\
  &= \lim\limits_{n\to \infty} \frac{1}{n} \paren*{ \log\abs{G'(x_n)} + \cdots + \log\abs{G'(x_n)} + \log\abs{h'(x_1)} } \\
  &= \lim\limits_{n\to \infty} \frac{1}{n} \paren*{ \log\abs{G'(x_n)} + \cdots + \log\abs{G'(x_n)} } \\
  &= \Lambda(y_1),
\end{align*} 
as desired.

\bibliographystyle{../../pjgsm}
\bibliography{../../bibliography/thesis}

\end{document}
