\documentclass[10pt,twoside,draft]{book}
\usepackage{../../thesis}
\graphicspath{ {../../images/} }
\usetikzlibrary{arrows,positioning} 
\tikzset{
    %Define standard arrow tip
  >=stealth',
    %Define style for boxes
  box/.style={
    rectangle,
    rounded corners,
    draw=black, very thick,
    text width=6.5em,
    minimum height=2em,
  text centered},
    % Define arrow style
  imply/.style={
    ->,
    thick,
    shorten <=2pt,
  shorten >=2pt},
  both/.style={
    <->,
    thick,
    shorten <=2pt,
  shorten >=2pt,},
  induced/.style={
    dotted,
    ->,
    thick,
    shorten <=2pt,
  shorten >=2pt},
  maybe/.style={
    ->, 
    -o,
    thick,
    shorten <=2pt,
  shorten >=50pt},
  noimply/.style={
    ->, %inden[
    -],
    thick,
    shorten <=2pt,
  shorten >=60pt}
}


\makeindex

\begin{document}
\label{chap:comparison}
\chapter{Comparisons of Definitions}
We have seen four definitions 
\begin{enumerate}[(i)]
  \item Devaney (\dev)
  \item Li-Yorke (\liy)
  \item Block-Coppel (\blcp)
  \item Positive topological entropy (\pte)
\end{enumerate}
and their variants
\begin{itemize}
  \item Wiggins's (similar to Devaney) (\wig)
  \item Martelli (equivalent to Wiggins's in $\R^n$)
  \item Marotto (implies Li-Yorke. Requires differentiability).
\end{itemize}
In this chapter, we discuss the relations between the definitions (i)$\sim$(iv) and Wiggins's definition.
Throughout this chapter, we consider the dynamical system $(X,F)$, where $X$ is a compact metric space, and $F$ is a continuous mapping.
We first discuss a special case when $X$ is a compact interval, then proceed to the general case.
%%%

\section{Compact Interval}
Let $I \subseteq \R$ be a compact interval.
In this section, we consider the case when $X = I$.
We show that \dev, \wig, \blcp, and \pte are equivalent conditions, and any of the three implies \liy.
\begin{figure}[ht]
  \centering
  \begin{tikzpicture}[node distance=1cm, auto,]
 %nodes
    \node[box] (ly) {Li-Yorke};
    \node[box, inner sep=5pt,above=1.0cm of ly] (bc) {Block-Coppel};
    \node[box, inner sep=5pt,left=1.5cm of bc] (dev) {Devaney};
    \node[box, inner sep=5pt,right=1.5cm of bc] (ent) {Entropy};
    \node[box, inner sep=5pt,above=1.0cm of dev] (wig) {Wiggins};
 %edges
    \draw[imply] (dev.south) to node[bend right=20] {} (ly.west);  %this guy doesn't bend
    \draw[imply] (ent) to[bend left=20] node {} (ly); 
    \draw[imply] (bc) to node {} (ly); 
    \draw[both] (dev) to[bend left=30] node {} (ent); 
    \draw[both] (dev) to node {} (bc); 
    \draw[both] (dev) to node {} (wig); 
    \draw[both] (bc) to node {} (ent); 
  \end{tikzpicture}
  \label{fig:chaos-interval}
  \caption{
    Relations between the five definitions for the compact interval case.
    The three definitions in the top row are all equivalent.
    Li-Yorke's definition is strictly weaker than the other three.
  }
\end{figure}

\subsubsection*{Devaney/Wiggins and Block-Coppel}
We already mentioned in Chapter~\ref{chap:devaney} that \dpp and \tt imply \sdic in a compact metric space.
Moreover, on a compact interval, topological transitivity implies the existence of dense periodic points.
Thus, Devaney's and Wiggins's definitions are equivalent.
\begin{theorem}
  $(I,F)$ is chaotic in the sense of Devaney if and only if it is chaotic in the sense of Wiggins.
  \label{thm:dev-wig}
  \begin{proof}
    It follows from the fact that \tt implies \dpp on a compact interval.
    See \citet{silverman} for the proof.
  \end{proof}
\end{theorem}
Also, \dev and \blcp are equivalent.
\begin{theorem}
  $(I,F)$ is chaotic in Devaney's sense if and only if it is chaotic in Block-Coppel's sense.
  \begin{proof}
    See \citet{aulbach}.
    %The proof uses \citep[VI: Proposition 6]{blockcoppel} .
  \end{proof}
  \label{thm:dev-blcp}
\end{theorem}
%%%

\subsubsection*{Block-Coppel and Topological Entropy}
\blcp is equivalent to \pte.
\begin{theorem}
  $(I,F)$ is chaotic in the sense of Block-Coppel if and only if $F: I \to I$ has a positive topological entropy.
  \label{thm:ent-blcp}
  \begin{proof}
    See \citet[VII, Theorem 24]{blockcoppel}.
  \end{proof}
\end{theorem}
Hence, the four definitions are equivalent.
%%%

\subsubsection*{Block-Coppel and Li-Yorke}
\blcp implies \liy.
\begin{theorem}
  \citep{blockcoppel}
  If $(I,F)$ is chaotic in Block-Coppel's sense, then it is chaotic in Li-Yorke's sense.
  \label{thm:blcp-liy}
  \begin{proof}
    See \citet{blockcoppel} (VI; Proposition 27).
  \end{proof}
\end{theorem}
However, the converse of the previous theorem is not true in general.
\begin{theorem}
  \citep[p.136]{blockcoppel}
  Suppose $(I,F)$ is not chaotic in Block-Coppel's sense.
  If there exists an asymptotically periodic point in $I$, then $(I,F)$ is chaotic in Li-Yorke's sense.
  \label{thm:blcp-per-liy}
\end{theorem}
Also, the following is an example of a dynamical system that is \liy but not \blcp.
\begin{example}
  (Truncated tent map \citep{aulbach})
  The tent map is defined as
  \begin{equation*}
    T(x) = 
    \begin{cases}
      2x     &\mbox{ if } x \in [0,1/2] \\
      2 - 2x &\mbox{ if } x \in [1/2,1].
    \end{cases}
  \end{equation*}
  We define the \textit{truncated tent map} for $\lambda \in [0,1]$ as 
  \begin{equation*}
    T_\lambda(x) = \min\set{\lambda, T(x)}.
  \end{equation*}
  (Thus, a truncated tent map is the tent map truncated at the height $\lambda$.
  Also, $T_1 = T$.)

  \begin{figure}[th]
    \centering
    \begin{tikzpicture}[scale=3.2]
      \draw[->] (-0.2,0) -- (1.2,0) node[right] {$x$};
      \draw[->] (0,-0.2) -- (0,1.2) node[above] {$T(x)$};
      \foreach \x/\xtext in {0.5/\frac{1}{2}, 1/1}
      \draw[shift={(\x,0)}] (0pt,1pt) -- (0pt,-1pt) node[below] {$\xtext$};
      \foreach \y/\ytext in {0.5/\frac{1}{2}, 0.8/\frac{4}{5}, 1/1}
      \draw[shift={(0,\y)}] (1pt,0pt) -- (-1pt,0pt) node[left] {$\ytext$};

      \draw[domain=0:0.4] plot (\x,{2*\x}) node[below right] {};
      \draw[domain=0.4:0.5, dotted] plot (\x,{2*\x}) node[below right] {};
      \draw[domain=0.5:0.6, dotted] plot (\x,{2 - 2*\x}) node[below right] {};
      \draw[domain=0.6:1] plot (\x,{2 - 2*\x}) node[below right] {};
      \draw[domain=0.4:0.6] plot (\x,{0.80}) node[below right] {};
    \end{tikzpicture}
    \label{fig:tent-map}
    \caption{The truncated tent map, $T_\lambda(x)$, for $\lambda = 4/5$.}
  \end{figure}
  Define a partition of $[0,1]$ by two sets
  \begin{equation*}
    J_\lambda \ceq \brac{0, \frac{\lambda}{2}} \cup \brac{1 - \frac{\lambda}{2}, 1} 
    \quad\mbox{ and }\quad
    K_\lambda \ceq \paren{\frac{\lambda}{2}, 1- \frac{\lambda}{2}}.
  \end{equation*}
  For any $0 \leq \lambda \leq \gamma \leq 1$, $T_\lambda$ and $T_\gamma$ coincide on $J_\lambda$.
  Furthermore, a point in $J_\lambda$ is a periodic point of $T_\gamma$ if and only if it is a periodic point of $T_\lambda$.

  The (untruncated) tent map $T_1$ has a 3-periodic point (e.g. $\set{2/7, 4/7, 6/7}$), so by Sarkovskii's theorem, $T_1$ has $2^n$-periodic points for all $n \in \N$.
  For any $n \in \N$, define
  \begin{equation*}
    \lambda_n \ceq \min\setst{\lambda \in [0,1]}{\mbox{ a } 2^n\mbox{-periodic orbit is contained in } [0,\lambda]}.
  \end{equation*}
  $\lambda_n$ exists for each $n$ because, for each $m \in \N$, the number of $m$-periodic points of $T_1$ is finite.
  To see this, note that $T_1$ has at most 2 fixed points, $\itr{T_1}{2}$ has at most $2^2$, and so on. 
  We can inductively show that the number of fixed points for $\itr{T_1}{m}$ is less than or equal to $2^m$. (Diagram)
  Also note that $\lambda_n$ is a $2^n$-periodic point of $T_1$ by construction.

  Next, we show that 
  \begin{equation*}
    O_{T_1}(\lambda_n) \subseteq J_{\lambda_n}.
  \end{equation*}
We have $T_1(K_{\lambda_n}) = (\lambda_n, 1]$, but by definition of $\lambda_n$, $O_{T_1}(\lambda_n)$ is contained in $[0,\lambda_n]$.
It follows that no point of $O_{T_1}(\lambda_n)$ is in $K_{\lambda_n}$, which implies $O_{T_1}(\lambda_n) \subseteq J_{\lambda_n}$.
Since $T_1$ and $T_{\lambda_n}$ coincide on $J_{\lambda_n}$, $\lambda_n$ is also a $2^n$-periodic point of $T_{\lambda_n}$, and its orbit under iteration by $T_{\lambda_n}$ is also contained in $J_{\lambda_n}$.
Then, by Sarkovskii's theorem, $T_{\lambda_n}$ has a periodic points of periods
\begin{equation*}
  2^n, 2^{n-1}, \ldots, 2, 1.
\end{equation*}
Furthermore, we show that no periodic point of $T_{\lambda_n}$ has a period different from these.
%We first prove that there cannot be more than one periodic point in $\cl(K_{\lambda_n})$ by contradiction. %%% Unnecessary?
%Suppose there exist $a,b \in \cl(K_{\lambda_n})$ that are periodic points of $T_{\lambda_n}$.
%Since $T_{\lambda_n}(a) = T_{\lambda_n}(b) = \lambda_n$, $a$ and $b$ are in the orbit of $\lambda_n$, so that the periods of $a$ and $b$ are $2^n$.
%We may suppose that $\itr{T}{k}(a) = b$ for some $0 < k < 2^n$ without loss of generality.
%Then,
%\begin{equation*}
%  \itr{T}{k}(a) = b 
%  \Rar 
%  \itr{T}{k+1}(a) = T(b)
%  \Rar 
%  \itr{T}{k}(\lambda_n) = \lambda_n,
%\end{equation*}
%which implies that $\lambda_n$ has period less than $2^n$, a contradiction.
To prove the claim, suppose on the contrary that $T_{\lambda_n}$ has a $m$-periodic point $p$, where $m$ is an integer that comes before $2^n$ in the Sarkovskii's ordering.
Denote $O_{T_{\lambda_n}}(p)$, the orbit of $p$, by $M$.
Since $T_{\lambda_n}$ can only have periodic points of period $2^n$ in $K_{\lambda_n}$, no point in $M$ is contained in $K_{\lambda_n}$, which means that
\begin{equation*}
  M \subseteq J_{\lambda_n}.
\end{equation*}
Now, define 
\begin{equation*}
  \rho \ceq \max M.
\end{equation*}
Since the period of $\rho$ is not $2^n$, it cannot be $\lambda_n$.
It follows that
\begin{equation*}
  \rho < \lambda_n.
\end{equation*}
Then, by Sarkovskii's theorem, there exists a $2^n$-periodic point in $[0,\rho] \subsetneq [0, \lambda_n]$.
This contradicts the fact that $\lambda_n$ is the minimum of such numbers.

To conclude the construction of the mapping, we show that
\begin{equation*}
  \lambda^* \ceq \lim\limits_{n \to \infty} \lambda_n
\end{equation*}
exists.
To see that the limit exists, first note that $\seq{\lambda_n}{\infty}{n = 0}$ is a strictly increasing sequence.
Also, $\seq{\lambda_n}{\infty}{n = 0}$ is bounded above by 1.
It follows that the sequence converges.
(\citet{misiurewicz1} showed that $\lambda^* = 0.8249080 \cdots$.)

Next, we show that $T_{\lambda^*}$ is not chaotic in the sense of Block-Coppel.
The proof is by contradiction.
By Theorem~\ref{thm:blcpII}, $T_{\lambda^*}$, a Block-Coppel chaotic map, has a periodic point whose period is not a power of 2.
So suppose $p \in [0,1]$ is a periodic point of period $q\cdot 2^n$ for some $n \geq 0$ and an odd integer $q > 1$.
Let 
\begin{equation*}
  \rho \ceq \max O(p).
\end{equation*}
If $\rho < \lambda^*$, then there exists a positive integer $m$ such that $\lambda_m > \rho$.
Note that $p$ is also a $q\cdot 2^n$-periodic point of $T_{\lambda_m}$.
However, there cannot be a periodic point whose period is not a power of two, so this is a contradiction.
If, on the other hand, $\rho = \lambda^*$, then by Sarkovskii's theorem, $T_{\lambda^*}$ has a periodic point $p'$ of period $(q+2)\cdot 2^n$.
Since $\max O(p') < \lambda^*$, this case reduces to the first case, which leads to a contradiction.
Thus, $T_{\lambda_m}$ is not Block-Coppel chaotic.

Furthermore, \citet[p.146]{blockcoppel} showed that $T_{\lambda^*}$ is chaotic in the sense of Li-Yorke using Theorem~\ref{thm:blcp-per-liy}. $\square$
\label{eg:counterexample}
\end{example}

%%%
\subsubsection*{Result}
It follows from Theorem~\ref{thm:dev-blcp} and Theorem~\ref{thm:ent-blcp} that \dev and \pte are also equivalent.
Hence, \dev, \wig, \blcp, and \pte are equivalent to each other on a compact interval.
Each of the four equivalent definitions implies \liy, but \liy does not imply others.
These results can be proved through other means.
\citet{smital} constructed an example of a Li-Yorke chaotic map that has zero topological entropy.
\citet{omegachaos} showed the equivalence of \dev and \pte, via his definition of chaos called $\omega$-chaos.
Also, some of these results are consequences of results in the general compact metric case.
%%%

\section{Compact Metric Space}
The result in this more general case is more varied than the compact interval case.
For instance, \dev, \wig, \blcp, and \pte are no longer equivalent to each other.
Also, there are questions that remain unanswered.
%Moreover, neither of them implies the others.
\begin{figure}[p]
  \centering
  \begin{tikzpicture}[node distance=1cm, auto]
 %nodes
    \node[] (dummy) {};
    \node[box, inner sep=5pt,below=2.0cm of dummy] (ly) {Li-Yorke};
    \node[box, inner sep=5pt,left=2.0cm of dummy] (dev) {Devaney};
    \node[box, inner sep=5pt,above=2.0cm of dummy] (bc) {Block-Coppel};
    \node[box, inner sep=5pt,right=2.0cm of dummy] (entropy) {Entropy};
    \node[below=1.0cm of ly] (dummy2) {};
 %edges
    \draw[imply] (entropy)[bend left=15] to node {} (ly.east); 
    \draw[imply] (dev)[bend right=15] to node {} (ly.west); 
    \draw[imply] (bc.east) to [bend left=15] node {} (entropy); 
    \draw[induced] (bc)[bend right=25] to node {} (ly); 
    \draw[maybe] (entropy)[] to node {?} (bc.east); 
    \draw[maybe] (entropy)[bend right=5] to node {?} (dev); 
    %\draw[noinduced] (dev)[bend left=15] to node {} (bc.west); 
    %\draw[noimply] (ly)[] to node {} (entropy); 
    %\draw[noimply] (ly)[] to node {} (dev); 
    %\draw[noimply] (ly)[] to node {} (bc); 
    %\draw[noimply] (bc)[] to node {} (dev); 
    %\draw[noimply] (dev)[bend right=15] to node {} (entropy); 
  \end{tikzpicture}
  \begin{tikzpicture}[node distance=1cm, auto]
 %nodes
    \node[] (dummy) {};
    \node[box, inner sep=5pt,below=2.0cm of dummy] (ly) {Li-Yorke};
    \node[box, inner sep=5pt,left=2.0cm of dummy] (dev) {Devaney};
    \node[box, inner sep=5pt,left=1.0cm of dev] (wig) {Wiggins};
    \node[box, inner sep=5pt,above=2.0cm of dummy] (bc) {Block-Coppel};
 %edges
    \draw[imply] (dev)[bend right=15] to node {} (ly.west); 
    \draw[imply] (dev)[] to node {} (wig.east); 
    \draw[imply] (bc)[bend right=15] to node {} (wig.north); 
    \draw[imply] (bc)[] to node {} (ly); 
    %\draw[noinduced] (dev)[bend left=15] to node {} (bc.west); 
    %\draw[noimply] (ly)[] to node {} (entropy); 
    %\draw[noimply] (ly)[] to node {} (dev); 
    %\draw[noimply] (ly)[] to node {} (bc); 
    %\draw[noimply] (bc)[] to node {} (dev); 
    %\draw[noimply] (wig)[bend right=15] to node {} (ly); 
    %\draw[noimply] (dev)[bend right=15] to node {} (entropy); 
  \end{tikzpicture}
  \label{fig:chaos-metric}
  \caption{
    Relations between definitions for a compact metric space.
    An absence of an arrow means that the particular implication is not necessarily true.
    An arrow with a circular head (with a question mark) means it is an open question.
    An dotted arrow indicates a result that is not directly shown, but induced by other implications.
  }
\end{figure}
%%%


\subsubsection*{Topological Entropy and Li-Yorke}
As we have seen in the compact interval case that \liy does not imply $\pte$.
The converse still holds in this general setting.
\begin{theorem}
  If $F: X \to X$ has positive topological entropy, then it is chaotic in Li-Yorke's sense.
  \label{thm:ent-liy}
  \begin{proof}
    See \citet{blanchard}.
  \end{proof}
\end{theorem}
%%%

\subsubsection*{Block-Coppel and Topological Entropy}
\begin{theorem}
  If $(X,F)$ is chaotic in Block-Coppel's sense, then $F$ has positive topological entropy.
  \begin{proof}
    $(X,F)$ is semi-conjugate to the full one-sided shift $(\sigma, \Sigma)$, so that by Proposition~\ref{thm:ent-conj}, $h(\sigma) \leq h(F)$.
    Since $h(\sigma)$ is positive, we conclude that $h(F)$ is also positive.
  \end{proof}
  \label{thm:blcp-ent}
\end{theorem}
%
However, we do not know whether \pte implies \blcp.

%%%


\subsubsection*{Devaney and Topological Entropy}
\citet{glasner} provides an example of a symbolic system that is \dev but has zero entropy.
Hence, \dev does not imply \pte.
On the other hand, we do not know whether \pte implies \dev.
We have some related results, however.
\citet{glasner} proved that if a system is transitive, has no isolated point, and has positive topological entropy, then it is sensitive to initial conditions.
Thus, if we can prove that \pte implies the existence of a closed, invariant subset $Y \subseteq X$ on which $F$ is topologically transitive, then \pte implies \wig.
On the other hand, in order for \pte to imply \dev, \pte mush imply both \tt and \dpp.
While it seems plausible that topological transitivity follows from positive topological entropy, whether \pte implies the existence of dense periodic points is questionable.
%%%

\subsubsection*{Block-Coppel and Devaney/Wiggins}
\citet{aulbach} constructed an example to show that \blcp does not imply \dev.
\begin{example}
  \citep{aulbach} 
  Define $\tau: \Sigma \to \Sigma$ as the addition by $(1, 0, 0, \ldots) \in \Sigma$:
  \begin{equation*}
    \tau: s \mapsto s + (1, 0, 0, \ldots),
  \end{equation*}
  where the addition of two members of $\Sigma$ is defined in the same manner as the addition of two binary numbers.
  For example, 
  \begin{align*}
    (1, 0, 0, \ldots) + (1, 0, 0, \ldots) &= (0, 1, 0, \ldots) \\
    (0, 1, 0, \ldots) + (1, 0, 0, \ldots) &= (1, 1, 0, \ldots) \\
    (1, 1, 0, \ldots) + (1, 0, 0, \ldots) &= (0, 0, 1, \ldots).
  \end{align*}
  Also, we define $\tau(1,1,1, \ldots) \ceq (0,0,0, \ldots)$.
  It is easy to see that $\tau: \Sigma \to \Sigma$ is a homeomorphism, and does not have a periodic point of any period.
  Now, consider the dynamical system $(\Sigma \times \Sigma, F)$, where $F$ is defined as
  \begin{equation*}
    F: (s, t) \mapsto (\sigma(s), \tau(t)).
  \end{equation*}
  $\Sigma$ is compact, so by Tychonoff's theorem, $\Sigma \times \Sigma$ is also compact.
  $(\Sigma \times \Sigma, F)$ is semi-conjugate to the one-sided full shift by $\pi_1$, the projection of the first component (i.e. $\pi_1 \circ F = \sigma \circ \pi_1$).
  Hence, the system is chaotic in Block-Coppel's sense.
  However, the system is not chaotic in Devaney's sense, because it doesn't have a periodic point.
  (Moreover, \citet{blockcoppel} proved that the system is also chaotic in Li-Yorke's sense.
  Therefore, \blcp together with \liy do not imply \dev.)
  $\square$
\end{example}
%%%
The system that we just introduced is not topologically transitive on $\Sigma \times \Sigma$.
However, it may still be possible to find a subset $Y$ of $\Sigma \times \Sigma$ such that $F_{|Y}$ is topologically transitive.
In fact, \citet{auslander} proved that it is indeed the case.
In addition, they showed that $F$ restricted to the subset is also sensitive to initial conditions; hence, \blcp implies \wig.

On the other hand, we can deduce that \dev does not imply \blcp from the results so far.
Suppose \dev implies \blcp.
Then \dev implies \pte; however, this is not necessarily the case, as shown in the comparison of \dev and \pte.
This is a contradiction. 
Hence, \dev does not imply \blcp.

\subsubsection*{Block-Coppel and Li-Yorke}
\blcp implies \pte, and \pte implies \liy; these implications induces the implication from \blcp to \liy.
However, we know that the converse is not true from the compact interval case.
Hence, \blcp implies \liy, but the converse is not true.

\subsubsection*{Devaney/Wiggins and Li-Yorke}
The following example is a system that is chaotic in Wiggins's sense but not in Li-Yorke's sense.
\begin{example}
  \citep{blanchard}
  This example employs a \textit{two-sided shift}, a symbolic dynamical system on the set of all bi-infinite sequences.
  As we mentioned in Chapter~\ref{chap:symbolic}, the basic structure of a two-sided shift is similar to an one-sided shift.
  We note here a few differences that we need for our example.
  Some of the results that we prove in this example are standard results for two-sided shifts.
  Let $S = \set{0,1}$ be the set of symbols.
  In a two-sided shift system, the space $\Sigma \equiv S^\Z$ consists of bi-infinite sequences, i.e. sequences of the form
  \begin{align*}
    \cdots s_{-n} \cdots s_{-2} s_{-1} s_0 s_1 s_2 \cdots s_n \cdots.
  \end{align*}
  We use the following metric on the space $\Sigma$ instead:
  \begin{equation*}
    \metric{a,b} = \sum\limits_{i = -\infty}^{\infty} \frac{\delta_i}{2^{\abs{i}}},
  \end{equation*}
  where
  \begin{equation*}
    \delta_i = 
    \begin{cases}
      &0 \quad \mbox{ if } a_i = b_i  \\
      &1 \quad \mbox{ otherwise.}
    \end{cases}
  \end{equation*}
  The shift map $\sigma$ for a two-sided shift is defined as
  \begin{equation*}
    \sigma: 
    \cdots s_{-n} \cdots s_{-2} s_{-1} * s_0 s_1 s_2 \cdots s_n \cdots
    \mapsto
    \cdots s_{-n} \cdots s_{-2} s_{-1} s_0 * s_1 s_2 \cdots s_n \cdots,
  \end{equation*}
  where $*$ is used to denote the 0th position.
  %This example uses a dynamical system commonly called the \textit{Sturmian system}.

  Let $S^1 \equiv [0, 1)$ be the unit circle, and let $R_\alpha: S^1 \to S^1$ be the rotation by an irrational number $\alpha \in S^1$. %]
  Recall that we showed in Chapter~\ref{chap:devaney} that the dynamical system $(S^1, R_\alpha)$ is topologically transitive, and has no periodic point.
  For each point $x \in S^1$, $\phi: S^1 \to \Sigma$ sends $x$ to $s \in \Sigma$ such that
  \begin{equation*}
    s = \cdots s_{-n} \cdots s_{-1} s_0 s_1 \cdots s_n \cdots,
  \end{equation*}
  where
  \begin{equation*}
    s_i = \begin{cases}
      0& \mbox{ if } \itr{R_\alpha}{i}(x) \in [0, \alpha)  \\ %]
      1& \mbox{ if } \itr{R_\alpha}{i}(x) \in [\alpha, 1). %]
    \end{cases}
  \end{equation*} %( (
  We show that $\phi$ is injective.
  First of all, we prove a lemma.
  %%%
  \begin{lemma}
    For each interval $I$ in $S^1$, there exists $n \in \N$ such that $0 \in \itr{R_\alpha}{n}(I)$.
    \label{lem:sturmian1}
    \begin{proof}
      We may assume that $I = (x,y)$, since if $(x,y)$ contains $0$, then $[x,y]$ also contains $0$.
      Then, let $z$ be the midpoint of the interval $(x,y)$, and $\eta \ceq \frac{\metric{x,y}}{2}$, so that $(z-\eta, z+\eta) = (x,y)$.
      Since $R_\alpha$ is transitive, there exists $m \geq 0$ such that 
      \begin{equation*}
        \itr{R_\alpha}{m}(z) \in \oball{\eta/2}{0}.
      \end{equation*}
      Therefore,
      \begin{equation*}
        \oball{\eta/2}{0} \subseteq \itr{R_\alpha}{m}(z-\eta, z+\eta) 
        = \itr{R_\alpha}{m}(x,y),
      \end{equation*}
      so that $0 \in \itr{R_\alpha}{m}(x,y)$.
    \end{proof}
  \end{lemma}
  %%%
  \begin{proposition}
    $\phi$ is injective.
    \begin{proof}
      Suppose $x \neq y \in S^1$.
      Let $I \ceq (x,y)$.
      By Lemma~\ref{lem:sturmian1}, there exists $n$ such that $0 \in \itr{R_\alpha}{n}(I)$.
      Then, $\phi(x)_n \neq \phi(y)_n$.
      Hence, $\phi(x) = \phi(y)$ implies $x = y$.
    \end{proof}
  \end{proposition}
  Now, let $X \ceq \phi(S^1)$.
  We consider the dynamical system $(\bar{X}, \sigma)$, where $\bar{X}$ is the closure of $X$.
  \begin{proposition}
    $\bar{X}$ is a compact subset of $\Sigma$, and is invariant under $\sigma$.
    (For $(\bar{X}, \sigma)$ to be a well-defined dynamical system, we need $\bar{X}$ to satisfy these conditions \citep[p.179]{lind}.)
    \begin{proof}
      Since $\bar{X}$, the closure of $X$, is a closed subset of a compact space $\Sigma$, $\bar{X}$ is also compact.
      It is clear by the construction of $X$ that $\sigma(X) = X = \itr{\sigma}{-1}(X)$.
      Suppose $s \in \bar{X}$.
      We may assume that there exists a sequence $\set{s_n} \subseteq X$ such that $\lim\limits_{n \to \infty} s_n = s$.
      For each $s_n$, there exists $\itr{\sigma}{-1}(s_n) \in X$.
      Since $\sigma$ is homeomorphic, we have 
      \begin{equation*}
        \lim\limits_{n \to \infty} \itr{\sigma}{-1}(s_n) 
        = \itr{\sigma}{-1}(s) 
        \in \cl(\itr{\sigma}{-1}(X)) 
        = \bar{X},
      \end{equation*}
      because $X$ is an invariant set of $\sigma$.
      It follows that $s \in \sigma(\bar{X})$; hence, $\bar{X} \subseteq \sigma(\bar{X})$.
      A similar argument shows that $\sigma(\bar{X}) \subseteq \bar{X}$.
      Hence, $\sigma(\bar{X}) = \bar{X}$.
    \end{proof}
  \end{proposition}
  Next, we prove that $(\bar{X}, \sigma)$ is transitive.
    %%%
  \begin{lemma}
    For each $x \in S^1$ and any $k \in \N$, there exists $\epsilon > 0$ such that $\oball{\epsilon}{x}$ contains a point $y$ such that 
    $\phi(x)_i = \phi(y)_i$ for each $0 \leq \abs{i} \leq k$.
    \label{lem:sturmian2}
    \begin{proof}
      It is easy to see that $\phi(x)_i \neq \phi(y)_i$ if and only if
      \begin{equation*}
      0 \in (\itr{R_\alpha}{i}(x), \itr{R_\alpha}{i}(y)] \mbox{ or } \alpha \in (\itr{R_\alpha}{i}(x), \itr{R_\alpha}{i}(y)]. %))
    \end{equation*}
    So for each $x \in S^1$, let $y = x + \delta$ for some $\delta > 0$.
  Also, let $I \equiv (x, y]$.
  If 
  \begin{equation*}
    0 \in R_\alpha(I) \mbox{ or } \alpha \in R_\alpha(I), 
  \end{equation*}
  then we can redefine $\delta$ to be a smaller value so that 
  \begin{equation*} 
    0, \alpha \not\in R_\alpha(I).
  \end{equation*}
  Inductively, we see that, for any $k \in \N$, if $\delta$ is small enough, then 
  \begin{equation*}
    0, \alpha \not\in \itr{R_\alpha}{k}(I)
  \end{equation*}
  so that
  \begin{equation*}
    \phi(x)_i = \phi(x+\delta)_i \mbox{ for } 0 \leq i \leq k.
  \end{equation*}
  Repeat the same process for backward iterations of $R_\alpha$ to obtain a sufficiently small $\delta$ such that
  \begin{equation*}
    \phi(x)_i = \phi(x+\delta)_i \mbox{ for } -k \leq i \leq 0.
  \end{equation*}
  This completes the proof.
\end{proof}
  \end{lemma}
  %%%

  \begin{proposition}
    Let $x,y$ be members of $\Sigma$.
    If $x_i = y_i$ for each $0 \leq \abs{i} \leq k$, then $\metric{x,y} \leq 2^{-(k-1)}$.
    \label{prop:two-sided}
    \begin{proof}
      $\metric{x,y}$ achieves its upper bound when $x_i \neq y_i$ for each $\abs{i} \geq k+1$.
      \begin{align*}
        \metric{x,y}
        &\leq 2 \cdot \sum\limits_{i = k+1}^{\infty} \frac{1}{2^{\abs{i}}} \\
        &\leq 2 \cdot \paren{ \sum\limits_{i = 0}^{\infty} \frac{1}{2^{\abs{i}}} - \sum\limits_{i = 0}^{k} \frac{1}{2^{\abs{i}}}} \\
        &= 2 \cdot (2 - (2 - \frac{1}{2^k})) \\
        &= 2^{-(k-1)}.
      \end{align*}
    \end{proof}
  \end{proposition}
  %%%
  \begin{proposition}
    Let $F: X \to X$ be a continuous map, $Y$ be a subset of $X$.
    If $F$ restricted to $Y$ is topologically transitive, then $F: \cl(Y) \to \cl(Y)$ is also transitive.
    \label{prop:closure-transitivity}
    \begin{proof}
      We only need to prove that $F$ is transitive for the boundary points of $Y$.
      For any point $x$ on the boundary of $Y$, and for each neighborhood $N$ of $x$, $N$ contains a point in $Y$.
      Call that point $y$.
      Choose $\epsilon$ small enough that $\oball{\epsilon}{y}$ is contained in $N \cap Y$.
      Applying the transitivity property of $F$ to the open neighborhood gives us the desired result.
    \end{proof}
  \end{proposition}
  %%%
  \begin{proposition}
    $(\bar{X}, \sigma)$ is transitive.
    \begin{proof}
      By Proposition~\ref{prop:closure-transitivity}, we only need to show that $\sigma$ restricted to $X$, the interior of $\bar{X}$, is transitive.
      Fix $s, t \in X$, $\epsilon > 0$, and $\delta > 0$.
      We may assume that $\epsilon = 2^{-j}$ and $\delta = 2^{-k}$ for some $j,k \in \N$.
      We show that there exists a point $t'$ in $\oball{\epsilon}{t}$ such that, for some $n$, $\itr{\sigma}{n}(t') \in \oball{\delta}{s}$.
      By Lemma~\ref{lem:sturmian2}, if $\eta > 0$ is small enough, then for any $x \in S^1$, we have $t \ceq \phi(x)$ and $t' \ceq \phi(x + \eta)$ satisfy $t_i = t'_i$ for $0 \leq \abs{i} \leq j$.
      Furthermore, by Proposition~\ref{prop:two-sided}, $t'$ satisfies
      \begin{equation*}
        \metric{t, t'} \leq 2^{-(j-1)} < \epsilon.
      \end{equation*}
      Hence, $t'$ is in $\oball{\epsilon}{t}$.
      Now, let $y \ceq \phi^{-1}(t') \in S^1$.
      By the transitivity of $R_\alpha$, we may find $m \in \N$ such that
      \begin{equation*}
        \itr{R_\alpha}{m}(y) \in \oball{\eta}{\phi^{-1}(s)},
      \end{equation*}
      where $\eta > 0$ taken so small that by Lemma~\ref{lem:sturmian2}, we have 
      $\phi(\itr{R_\alpha}{m}(y))_i = s_i$ for $0 \leq \abs{i} \leq k$.
      Note that
      \begin{equation*}
        \phi(\itr{R_\alpha}{m}(y)) 
        = \itr{\sigma}{m}(\phi(y))
        \equiv \itr{\sigma}{m}(t).
      \end{equation*}
      Therefore, $t'_{m+i} = s_i$ for $0 \leq \abs{i} \leq k$.
      It follows that 
      \begin{equation*}
        \metric{\itr{\sigma}{m}(t'), s} \leq 2^{-(k-1)} < \delta.
      \end{equation*}
      This completes the proof.
    \end{proof}
  \end{proposition}
  %%%
  We are left to show that $(\bar{X}, \sigma)$ is sensitive to initial conditions in order to prove that the system is chaotic in Wiggins's sense.
  \begin{proposition}
    $(\bar{X}, \sigma)$ is sensitive to initial conditions.
    \begin{proof}
      Let $0 < \delta < 1$ be our separation constant.
      For any given $x \in S^1$, and $\epsilon > 0$, we can find $y \in \oball{\epsilon}{x}$ by Lemma~\ref{lem:sturmian2}.
      If $y \geq x$, then let $I \ceq (x,y)$; otherwise, let $I \ceq (y,x)$.
      By Lemma~\ref{lem:sturmian1}, there exists $m \in \N$ such that $\phi(x)_m \neq \phi(y)_m$.
      Therefore,
      \begin{equation*}
        \metric{\itr{\sigma}{m}(x), \itr{\sigma}{m}(y)} \geq 1 > \delta.
      \end{equation*}
    \end{proof}
  \end{proposition}
  %%%
  Our next goal is to show that $(\bar{X}, \sigma)$ is not chaotic in Li-Yorke's sense.
  To do so, we prove 
  \begin{equation*}
    \liminf\limits_{n \to \infty} (\itr{\sigma}{n}(s), \itr{\sigma}{n}(t)) > 0
  \end{equation*}
  for each $s, t \in \Sigma$.
  \begin{proposition}
    $(\bar{X}, \sigma)$ is not chaotic in Li-Yorke's sense.
    \begin{proof}
    By Lemma~\ref{lem:sturmian1}, we know that there exists $n$ such that $(\itr{R_\alpha}{n}(x), \itr{R_\alpha}{n}(y)]$ contains $0$. %)
  We extend the result and show that the interval $(\itr{R_\alpha}{n}(x), \itr{R_\alpha}{n}(y)]$ comes to contain $0$ at a certain frequency under iteration. %)
  In particular, we prove that the number of time steps between the two such occurences cannot be arbitrarily long.

  First, choose any $x,y \in S^1$. %[
  Let $b$ be the length of the interval $I \equiv (x,y]$. %)
  Because of Lemma~\ref{lem:sturmian1}, we may assume without loss of generality that $I$ contains $0$.
  First, consider the case $\alpha \leq b$.
  In this case, $I \cap R_\alpha(I)$, so as $I$ rotates around $S^1$ once, it covers all of $S^1$.
  Let $n$ be the least positive integer such that $n\alpha > 1$.
  Then, it takes no longer than $n$ iterations until $I$ contains $0$ again.

  Next, consider the other case $\alpha > b$.
  There exists a positive integer $m$ such that $m\alpha \leq b \;(\mod{1})$.
  Then, as in the first case, if we let $n$ be the least positive integer such that $nm\alpha > 1$, then $mn$ is the upper bound for the number of iterations until $\itr{R_\alpha}{m}(I)$ contains $0$ again.

  Now, consider $s \ceq \phi(x)$ and $t \ceq \phi(y)$.
  By the discussion so far, each block of length $mn$ contains at least one position $i$ such that $\phi(x)_{i} \neq \phi(y)_{i}$.
  It follows that the longest block for which $s$ and $t$ agree with each other cannot be longer than $2mn$.
  Hence,
  \begin{equation*}
    \metric{\itr{\sigma}{i}(x), \itr{\sigma}{i}(y)} 
    \geq 2 \cdot \frac{1}{2^{mn}}
    = 2^{1 - mn}.
  \end{equation*}
  Thus, $\liminf\limits_{i \to \infty} (\itr{\sigma}{i}(x), \itr{\sigma}{i}(y)) > 0$.
\end{proof}
  \end{proposition}
  %%%
  Finally, we show that the system does not have a periodic point.
  In particular, this implies that $\sigma$ does not have dense periodic points in $\bar{X}$.
  \begin{proposition}
    $\sigma$ does not have a periodic point in $\bar{X}$.
    \begin{proof}
      Proof by contradiction.
      Suppose $s \in X$ is a periodic point of period $p$, and let $x \ceq \phi^{-1}(s)$. 
      Without loss of generality, we may assume that $x \in [0, \alpha)$. %]
      Then, for each $k \in \N$, we have $\itr{R_\alpha}{kp}(x) \in [0, \alpha)$. %]
      However, this is impossible because $\itr{R_\alpha}{p}$ is an irrational rotation, and hence transitive.
    \end{proof}
  \end{proposition}
  %%%
  Hence, $(\sigma, \bar{X})$ is chaotic in Wiggins's sense, but not in Li-Yorke's sense.
  The last proposition implies that the system is not chaotic in Devaney's sense.
  $\square$
  \label{eg:blanchard}
\end{example}
%%%
As we have seen in the example above, \wig does not necessarily imply \liy.
However, \dev, which is stronger than \wig, implies \liy.
%%%
\begin{theorem}
  If $(X, F)$ is chaotic in Devaney's sense, then it is chaotic in Li-Yorke's sense.
  \begin{proof}
    See \citet{mai}.
    \citet{huang} also proved the same result.
  \end{proof}
  \label{thm:dev-liy}
\end{theorem}
More specifically, Mai proved that, for a complete metric space $X$ without an isolated point, if $F: X \to X$ is transitive and has a periodic point of any period, then $F$ is chaotic in Li-Yorke's sense.
Thus, even though \tt and \sdic do not imply Li-Yorke's chaos, as we saw in the previous example, \tt and the existence of a \textit{single} periodic point, which is significantly weaker than \dpp, imply chaos in Li-Yorke's sense.
This result may be instructive in evaluating whether \dpp in Devaney's definition is appropriate.
Also, we can make an interesting observation from the fact that \wig and \liy do not imply each other. 
Recall that, though it is not explicitly called sensitivity, Li-Yorke's definition uses its own version of sensitivity, $\limsup\limits_{n\to \infty} \metric{\itr{F}{n}(x), \itr{F}{n}(y)} > 0$, called the \textit{non-asymptotic relation}.
The non-asymptotic relation is analogous to, yet different from \sdic.
In the same manner, we can see that the other condition in Li-Yorke's definition, $\liminf\limits_{n\to \infty} \metric{\itr{F}{n}(x), \itr{F}{n}(y)} = 0$, called the \textit{proximal relation}, is similar to topological transitivity.
The result about \wig and \liy indicates that non-asymptotic and proximal relations are neither stronger nor weaker than \tt and \sdic, even though these conditions contain similar ideas.
%%%

\subsubsection*{Result}
As in the compact interval case, each of \dev, \blcp, and \pte implies \liy.
However, \wig is weaker than \dev, and does not imply \liy.
Moreover, the equivalences of definitions that we had for a compact interval do not hold (Figure~\ref{fig:chaos-metric}).
Also, there are open questions.
We do not know whether \pte implies \blcp or \dev.
We obtain the implication from \pte to \wig if \pte implies the existence of a subset $Y$ such that $F_{|Y}$ is topological transitive.
We may also be interested in looking at individual conditions in each definition.
For example, we may ask: does topological transitivity imply positive topological entropy?

We have a few more things to say about Devaney's definition.
We saw that a single periodic point in addition to \wig is a sufficient condition to imply \liy. 
This result suggests that the existence of dense periodic points may be unnecessarily strong.
As we mentioned in Chapter~\ref{chap:devaney}, \citet{glasner} proposed that requiring a dense set of periodic points is too restrictive, and suggested alternative conditions, one of which is the existence of an invariant measure that is positive on each non-empty open set.
Also, the contrast of \dev and \wig seen through their relations with \pte and \liy (Figure~\ref{fig:chaos-metric}; the second diagram) is noteworthy.
Because \dev is stronger, \dev implies \liy, but \wig does not.
By the same reason, \blcp implies \wig, but not \dev.

\bibliographystyle{../../bibliography/pjgsm}
\bibliography{../../bibliography/thesis}

\printindex

%\section{Conclusion}
%Work in progress
%We have concentrated on the comparisons of definitions, but the field of dynamical systems is more than just that.
%For example, one may be interested in finding what conditions imply \sdic.
%We proved that \dpp and \tt are sufficient conditions for \sdic, but can we use a weaker condition to derive \sdic?
%Are there other conditions that imply \sdic?
%Do topologically transitive maps have positive topological entropy?
%One could define chaos for measure-theoretic dynamics \citep{downarowicz} or differentiable dynamical systems \citep{ruellebook}.
%One could also study how complicated dynamics arise.
%For example, period doubling route to chaos and intermittency route to chaos \citep{sternbergbook}
%
%\subsection*{Some Open Problems}
%\begin{enumerate}
%  \item Steve's problem
%  \item Does topological transitivity imply positive topological entropy?
%\end{enumerate}


\end{document}
