\documentclass[12pt]{book}
\usepackage{../../thesis}

\begin{document}
\section{A note on sensitive dependence on initial conditions}
Some mappings qualify to sensitive dependence, even if it seems simple and predictable.

\begin{proposition}
  (Rotation by magnitude is chaotic)

  Let $D = \set{x \in \R^2: \norm{x} \leq 2}$, and $C_\rho = \set{x \in \R^2: \norm{x} = \rho}$.
  Define $F: D \to D$, a rotation in $C_\rho$, by
  \begin{equation*}
    F(x) = F(\rho, \theta) = (\rho, \theta + \rho).
  \end{equation*}
  For every $\rho \in (0,2]$, $C_\rho$ is invariant.

  We claim that the system has in $C_\rho$ sensitive dependence on initial conditions with $r_0 = \rho$.
  Let $x_0 = (\rho_0, \theta_0)$ and $d > 0$.
  Choose $n$ large enough that $\frac{\pi}{n} < d$ and $\rho_0 - \frac{\pi}{n} > 0$.
  Let $y_0 = (\rho_0 - \frac{\pi}{n}, \theta_0)$.
  Then,
  \begin{equation*}
    x_n = (\rho_0, \theta_0 + n\rho_0),\quad
    y_n = (\rho_0 - \frac{\pi}{n}, \theta_0 + n\rho_0 - \pi),
  \end{equation*}
  so
  \begin{equation*}
    \norm{x_0 - y_0} < d \quad \mbox{and} \quad \norm{y_n - x_n} = \paren*{\frac{\pi}{n}}^2 + \pi^2 > 2 > r_0.
  \end{equation*}
\end{proposition}
 
! Create a plot !

\bibliographystyle{pjgsm}
\bibliography{../../bibliography/thesis}

\section{Definiton of Chaos}
Sternberg Chap4
\begin{definition}
  (Dense Periodic Points) Let $f: I \rar I$. Define $P(f)$ to be the set of periodic points
  of the map $f$. We say that $f$ has dense periodic points if $P(f)$ is
  dense in $I$.
\end{definition}

\begin{proposition}
  (Topological Transitivity and Dense Periodic Points Imply Sensitivity to Initial Conditions)
  Let $f: X\rar X$ be a chaotic transformation. Then there is an $\delta > 0$ such that
  for any $x\in X$ and any open set $J$ containing at least $x$ and another point, there is
  a point $y\in J$ and $n\in \N$ with
  \begin{equation*}
    d(f^n(x),f^n(y)) > \delta.
  \end{equation*}
  \label{thm:sensitivity}
\end{proposition}
\begin{lemma}
  Let $f: X\rar X$ be a mapping with (at least) two distinct periodic orbits.
  Then there is a $c > 0$ such that for any $x\in X$ there exists a periodic
  point $p$ with the property that for each $k$
  \begin{equation*}
    d(x, f^k(p)) > c.
  \end{equation*}
\end{lemma}
\begin{proof}[Proof of the lemma]
  Let $r$ and $s$ be periodic points with distinct orbits. For each $k$ and $l$ we have
  \begin{equation*}
    d(f^k(r), f^l(s)) > 0.
  \end{equation*}
  Choose $c$ so that $2c < \min d(f^k(r),f^l(s)) > 0$. Then by the triangle inequality
  \begin{equation*}
    2c < d(f^k(r),f^l(s)) \leq d(f^k(r),x) + d(x,f^l(s))
  \end{equation*}
  for all $k$ and $l$. Now suppose that $d(f^k(r),x)<c$ and $d(x,f^l(s))<c$. The supposition
  violates the inequality just derived. Thus $d(f^k(r),x) \geq c$ or $d(x,f^l(s)) \geq c$ must hold.
\end{proof}
\begin{proof}[Proof of \ref{thm:sensitivity}]
  Let $\delta = c/4$. Let $x$ be any point of $X$ and $J$ any open set containing $x$.
  Since the periodic points of $f$ are dense (in $X$), we may introduce a periodic point
  $q$ of $f$ in
  \begin{equation*}
    U = J\cap B_\delta(x).
  \end{equation*}
  Let $n$ be the period of $q$.
  
  Let $p$ be a periodic point whose orbit is apart from $x$ by $4\delta$. Define
  \begin{equation*}
    W_i := B_i(f^i(p)) \text{ and } V := f^{-1}(W_1) \cap f^{-2}(W_2) \cap \cdots f^{-n}(W_n).
  \end{equation*}
  Since $f^i(p) \in W_i$, we see that $V$ is not empty. Note that $V$ is open.

  By the transitivity property, we can find a $z\in U$ and a positive integer $k$ such that
  $f^k(z) \in V$. Let $j$ be the smallest integer so that $k < nj$. That is,
  \begin{equation*}
    1 \leq nj - k \leq n.
  \end{equation*}
  Therefore,
  \begin{equation*}
    f^{nj - k} (V) = f^{nj - k} (f^k(z)) \in f^{nj-k}(V).
  \end{equation*}
  Also,
  \begin{align*}
     f^{nj - k}(V) &= f^{nj - k} (f^{-1}(W_1) \cap f^{-2}(W_2) \cap \cdots f^{-n}(W_n))
     &\subset f^{nj - k}(f^{-(nj - k)} W_{nj-k})
     &= W_{nj_k}
  \end{align*}
  In other words,
  \begin{equation*}
    d(f^{nj}(z), f^{nj - k}(p)) < \delta.
  \end{equation*}
  Furthermore, $f^{nj}(q) = q$, since $q$ is periodic with period $n$. Hence, 
  \begin{equation*}
    d(f^{nj}(q), f^{nj}(z)) = d(q, f^{nj}(z)) 
  \end{equation*}
  By the triangle inequality, we have
  \begin{equation*}
    d(x, f^{nj-k}(p)) \leq d(x,q) + d(q, f^{nj}(z)) + d(f^{nj}(z), f^{nj - k}(p)).
  \end{equation*}
  Finally, 
  \begin{equation*}
    d(f^{nj}(q), f^{nj}(z)) = d(q, f^{nj}(z)) \geq  d(x, f^{nj-k}(p)) - d(x,q) - d(f^{nj}(z), f^{nj - k}(p)) = 4\delta - \delta - \delta = 2\delta.
  \end{equation*}
  The inequality implies that $d(f^{nj}(x), f^{nj}(z)) \geq \delta$ or $d(f^{nj}(x), f^{nj}(q)) \geq \delta$ must hold.

\end{proof}


\end{document}

